---
title: "unknown codes investigation"
date: "`r lubridate::today()`"
output:
  html_document:
    code_folding: hide
---

```{r, rmd-setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(lubridate)
library(here)
```

# Overview

This report investigates which clinical codes are used in the TPP backend but not available in OpenCodelists.

The current version of the reports looks at the following tables:

- APCS_Der.Spell_Primary_Diagnosis (icd10)
- CodedEvent.CTV3Code (ctv3)
- CodedEvent_SNOMED.ConceptID (snomedct)

# Load and summarise data

- The date range was limited from 2015-01-01 to `r lubridate::today()`

```{r}
# Read data
df_codes <- read_csv(here("output/output.csv")) %>%
  mutate(month = as.Date(paste0(month, "-01"))) %>%
  mutate(code_unknown = factor(case_when(code_unknown ~ "Code unknown",
                                         !code_unknown ~ "Code known")),
         data_source = factor(data_source),
         coding_system = factor(coding_system)) %>%
  filter(month > as_date("2015-01-01") & month < today())
```

The following table shows the minimum and maximum number (`min_num`, `max_num`) and date (`min_date`, `max_date`) of known and unknown codes.

```{r}
# Count number of known and unknown codes for each coding system
df_codes %>%
  group_by(data_source, coding_system, code_unknown) %>%
  summarise(min_num = scales::comma(min(num)),
            max_num = scales::comma(max(num)),
            min_date = min(month),
            max_date = max(month)) %>%
  DT::datatable()
```


This table shows the count of unique known and unknown codes in TPP each table.

```{r}

df_codes %>%
  group_by(data_source, coding_system, code_unknown) %>%
  select(-c(month, num)) %>%
  distinct() %>%
  summarise(n_unique_unknown_codes = scales::comma(n())) %>%
  DT::datatable()

```

# Visualising monthly changes 

## Sum of known and unknown codes

```{r fig.width=9, fig.height=6}

# First, calculate sum of number of known/unknown codes used in TPP backend by month
df_codes_sum <- df_codes %>%
  group_by(data_source, coding_system, code_unknown, month) %>%
  mutate(code_sum_num = sum(num, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-num, -code) %>%
  distinct()

# Create plot
df_codes_sum %>%
    ggplot2::ggplot(ggplot2::aes(
      x = month,
      y = code_sum_num,
      colour = coding_system
    )) +
    ggplot2::geom_line(
      size = .7,
      alpha = 0.3
    ) +
    ggplot2::geom_point(
      size = 1,
      alpha = 0.9
    ) +
    ggplot2::scale_x_date(
      date_breaks = "6 months",
      labels = scales::label_date_short()
    ) +
    ggplot2::theme(text = ggplot2::element_text(size = 10)) +
    ggplot2::facet_wrap(~code_unknown, scales = "free_y", ncol = 1) +
    ggplot2::scale_y_continuous(labels = scales::comma)

```

## Top 5 unknown codes per coding system 

```{r fig.width=9, fig.height=12}

# Select top 5 unknown codes
df_codes_slice_max5 <- df_codes %>%
  filter(code_unknown == "Code unknown") %>%
  group_by(data_source, coding_system, code) %>%
  mutate(code_sum = sum(num, na.rm = TRUE)) %>%
  group_by(data_source, coding_system) %>%
  slice_max(code_sum, n = 5)

# Create individual plots for top 5 unknown
df_codes_slice_max5 %>%
  ggplot2::ggplot(ggplot2::aes(
      x = month,
      y = num,
      colour = coding_system
    )) +
    ggplot2::geom_line(
      size = .7,
      alpha = 0.3
    ) +
    ggplot2::geom_point(
      size = 1,
      alpha = 0.9
    ) +
    ggplot2::scale_x_date(
      date_breaks = "6 months",
      labels = scales::label_date_short()
    ) +
    ggplot2::theme(text = ggplot2::element_text(size = 10)) +
    ggplot2::facet_wrap(~code, scales = "free_y", ncol = 1) +
    ggplot2::scale_y_continuous(labels = scales::comma)

```

# Total count of distinct unknown codes 

The following table presents the sum of each unknown code used in TPP between 2015-01-01 to `r lubridate::today()`.
Note that codes were removed from this analysis if there were less than 10 occurences in a given month.

```{r}

df_codes %>%
  filter(code_unknown == "Code unknown") %>%
  group_by(data_source, coding_system, code) %>%
  mutate(code_sum = sum(num, na.rm = TRUE)) %>%
  ungroup() %>%
  select(code, coding_system, code_sum) %>%
  distinct() %>%
  arrange(-code_sum) %>%
  DT::datatable(filter = 'top',
                options = list(pageLength = 20))
```
